### 2022-05-24 Overview.Rmd
initial commit

---
title: "overview"
author: "Justin Pomeranz"
date: "5/24/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ABT
### attempt 1
Size spectra are fundamental descriptions of communities and are being used increasingly to assess ecological health. Numerous methods exist for estimating size spectra, each of which introduces its own unique systematic biases making direct comparisons of results difficult. But no studies have been peformed to examine if different methodologies are able to detect differences in size spectra parameters, or if the *relative* differences within a method are consistent. Therefore, we set out to answer this question to guide future research, as well as to leverage previously published results to synthesize knowledge of how size spectra relationships respond across studies.

### attempt 2
Community body size distributions are a fundamental characteristic of natural communities related to energy availability and trophic transfer efficiency. Numerous methodologies for estimating size spectra relationships have been proposed and compared, with maximum likelihood estimates generally being the most accurate. However, to date, no study has been conducted to assess which methodologies are able to detect known differences in size spectra parameters (slope or exponent). The present study addresses this shortfall by analyzing data using common methods on data simulated from know distributions.

## Motivation

Size spectra relationships describe the negative pattern of abundance with increasing body size. Communities across habitats, biomes, and scales have shown remarkably similar patterns, leading to body size distributions being proposed as a "universal" relationship (Pethcey and Belgrano 2010). Furthermore, deviations from this expected pattern have been demonstrated in communities which have experienced extreme anthropogenic stress, including from overfishing marine communities in the North Sea (Blanchard/Jennings (always forget which one)), and stream macroinvertebrate communities from acid mine drainage in New Zealand (Pomeranz et al. 2018). The generally consistent pattern in undisturbed communities and response to stressors of size spectra is increasing interest in the use of size spectra as a universal indicator in ecology.

Numerous methodologies have been proposed for estimating size spectra relationships. Traditionally, body sizes (M) have been grouped together in bins, and an ordinary least squares regression on the abundance (N) within each bin has been conducted in log-log space to estimate the slope (exponent) of decreasing N in response to increasing M. Many binning methods have been proposed and used previously in the literature. However, numerous studies have shown that binning methods introduce systematic error, and are generally do not provide accurate estimates of the slope in log-log space, or the exponent in real space. Maximum likelihood estimates of the exponent describing the power law have generally been shown to provide the most accurate estimate.

Despite the increasing consensus of MLE analyses, no study has yet been conducted to assess which method, if any, can accurately detect differences in parameters describing size spectra relationships. Knowing which methods are able to detect differences is an important goal for ecologists moving forward, particularly with the increasing interest in using size spectra relationships to characterize communities. Additionally, having an understanding of how previous estimates of size spectra parameters vary based on methodologies will allow future studies to leverage previously published results. Because of the variation in estimates based on methodologies, it is currently not possible to conduct meta-analyses synthesizing previously published data.

## methods

1. Simulate data
  + sizeSpectra::rPLB(n = 1000, b, xmin, xmax)
  + b = c(-1.5, -2, -2.5); b = c(-1.9, -2, -2.1)
  + xmin and xmax based on macroinvertebrate sizes
  + Also use Edwards values to test sensitivity
  + Simulate from other distributions? rlnorm(), rweibull() etc.?
2. Analyze simulated data
  + MLE methods from Edwards et al. 2018
  + log2 bins - Pomeranz, Blanchard, etc
  + ELB - 6, 8, 10 bins. Woodward's group generally uses 6 equal bins
  + Others from Edwards' 8 methods?
3. assessment
  + which methods provide best estimate of *real* slopes?
  + what is *relative* difference in estimates? i.e. do all methods result in a relative change of 0.5 between data "groups"?