### 2022-10-24 update_oct_6.Rmd
rmd oct 6 update

---
title: "Manuscript: Detecting differences in Size Spectra"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "jpz"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Quick update post discussion on 10/6

It seemed to me that there were three main points to cover based on our discussion.
1. How do $\lambda$ estimates change across known "shallow" to "steep" values?
2. How do our estimates of the relationships of $\lambda$ differ across "lambda windows"?
3. how do our estimates perform when we have a strong or weak relationship (i.e. $\beta_1$) across a gradient?

# 1 $\lambda$ estimates

![lambda estimates by method from sampling from random distributions of steep (lambda = -2.5) to shallow (lambda = -0.5) exponents](figures/est_lambda_est_b_density.png)

**NOTE** the difference in x-axis scale. 
The MLE (blue) distribution is nearly always centered on the known value of lambda exponent (dashed line, and facet title).

For reference, here is the plot without allown the x-axis to vary. I think it's really hard to see anyhting in this, but the varying x-axis above is also hard to intuit, so not sure which to use. 
![Same plot but with "fixed" x-scale](figures/est_lambda_est_b_density_fixed_x.png)

![Plot of the known lambda (wrong x-axis label) and the lambda estimate from different methods. The dashed line is the 1:1 line, and almost perfectly matches the MLE (Blue) estimate. The two binning method cross the 1:1 line, meaning they systematically underestimate more positive (shallow) lambdas, and over estimate more negative (steeper) lambdas.](figures/known_est_b_line.png)

This figure shows that methods are comparble when lambda = ~-1, but are systematically biased at extreme lambda values. Also note that the theoretical exponent for abundance spectra are $\lambda = 2$, and there is considerable variation at this point in the figure.

# 2 lambda "windows"

Because of the markedly different performance of the two binning methods at steep and shallow values of lambdas, we first investigated how well the methods capture a known relationship of beta = -0.5 across a hypothetical environmental gradient. The lambda "windows" are steep (-2.5, -1.5), medium (-2, -1) and shallow (-1.5, -0.5).

![Distribution of relationship estimates (beta_1) in three different "windows" of lambda values. The dashed vertical line is the known relationship value of -0.5.  ](figures/lambda_angle_plot.png)

The distribution of relationship estimates from the MLE method (blue) is approximately centered at the known relationship value. The distribution is widest in the "steep" window, but is still markedly narrower than the binning methods. In addition to having much wider distributions, the binning methods systematically overestimate the known relationship.

# 3 varying the known relationship

Finally, we were interested in finding if the three methods performed differently with varying values of the known relaitonship. Here, we present estimates

![regressions for three different known relationship values: beta = 0, 0.25, and 0.5](figures/vary_beta_plot.png)

Finally, we can also plot the distribution of relationship estimates

![Distribution of relationship estimates (beta_1) when estimating from different known relationships](figures/vary_beta_density_plot.png)
All the methods perform well when no relationship is present. The performance of the binning methods (median values) declines, with relationships being systematically over estimated. Likewise, the width of the distributions increases with stronger relationships across a hypothetical gradient. However, the distribution of estimates from the MLE method is nearly always centered at the known value, and has markedly narrower distributions.